---
title: "Appendix S2: Ecological drivers of parasite aggregation: a meta-analysis of adherence to the 20/80 rule"
author: "E. Sherrard-Smith, J. F. Stephenson, E. Gillingham, R. Brackley, D. Shaw, S. E. Perkins"
date: "26 March 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This supplement provides the script for the data preparation and calculation of *k*, Gini coefficient, Poulin's D and *t<sub>20*. The script and output of the analyses of the ecological factors potentially important in
determining parasite aggregation in host populations. The data used for this analysis are in the supplementary
file ‘SherrardSmithetalDATA.csv’ (Appendix S1; Table S1). 



## A step by step run through of the data analysis for 20:80 #####

We explore whether parasite populations fit the '20:80' distribution where 80% of the parasites are contained within 20% of the hosts. First, we explore real data on this relationship and then compare these 
distributions to those that are generated using the mean, N of hosts, and a measure of variance.


###Step 1
```{r}
library(MASS)
library(reshape)
```

###Example Using Real Data; supplied by Darren Shaw September 2013
```{r}
CnodosMarWilliams63 <- rep(seq(0,5,1),c(496,58,18,7,2,4))
```

First data is summarised to give the appropriate information on mean, variance, host and parasite number for the comparison of these real distributions with those that are estimated.

```{r}
data<-sort(CnodosMarWilliams63,decreasing=TRUE)
	data[1:200]
	sampmean<-mean(data)
	sampmean               ###3.446
		a<-numeric(10000)
		for(i in 1:10000) a[i]<-mean(sample(data,replace=T),na.rm=T)
		quantile(a,c(0.025,0.975))
	sumNpar<-sum(data)
	sumNpar
	N<-length(data)
	N
	var<-var(data)
	var
	k<-(sampmean^2-(var/N))/(var-sampmean)
	k
```


Generate the distribution of parasites in a given proportion of the hosts Using the real data distributions
```{r}
real2<-data.frame(data)

##distribution of parasites in the sample population
real3<-data.frame(sort(real2[,1],decreasing=TRUE))

samp<-length(real3[,1])
nmk<-numeric(samp)
mk<-matrix(data = NA, nrow = samp, ncol = 1) 

for (j in 1:samp) ## a nested loop which counts through the number of hosts
  {
            addup<-sum(real3[1:j,1])             ## cumulative sum of parasites 
     if (sum(real3)>0){nmk[j]<-addup/sum(real3)} ## if there are ANY parasites 
                                                 ## this gives the cumulative 
                                                 ## proportion transmission per host
            else {nmk[j]<-0 }                    ## if there are no parasites pop 
                                                 ## stops a problem due to division  
                                                 ## by 0
  }

mk[,1]<-nmk                                      ## puts nmk into matrix mk in column i
raw<-data.frame(mk)
props<-c(0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,0.5,
         0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99)
#the proportions we will investigate
```


```{r}
names<-paste("t",props,sep="")#names for each list of proportions
tvals<-matrix(NA,nrow=dim(raw)[2],ncol=length(props))# a matrix to store the tx values 
                                                     # rows = studies,  
                                                     # cols = tx values


for (j in 1:length(props))# first loop cycles through desired tx values 
                          # contained in props above
{
  for (i in 1:dim(raw)[2]) # second loop cycles through each column of the data 
                           # (which is a vector of proportions of parasites in each  
                           # host from most to least infected)
  {
    vec<-raw[,i]             # make a vector from the column
    vec<-na.omit(vec)        # take out NAs after values reach 1 (it's a ragged array)
    N<-length(vec)           # number of hosts = number of rows
    Nx<-round(props[j]*N)    # gets the row which is x of the way down the column
    if (Nx==0) tvals[i,j]<-0 # if it rounds Nx to 0, there is no such thing as 
                             # a zeroth row, so make tx 0
    else {tx<-vec[Nx]        # gets the value in that column - the proportion of  
                             # parasites captured in the x most infected hosts
    tvals[i,j]<-tx           # rowi=study index colj=tx index
    
    }
  }
  
}
t.vals_Observed<-t(tvals)
```

###Step 2 
###Generate the estimated distributions from the real data summary information; mean, N, var  
```{r}
dmean = 0.2444444                   #This is the sample mean in study 'Aa'
samp = 585                          #This is the number of hosts (N) in study 'Aa'
k = 0.2335546                       #This is the k estimate from the mean var and N 
                                    #in study 'Aa'

mk<-matrix(data = NA, nrow = samp, ncol = 100)  # creates a matrix with
                                                # rows equal to sample size and  
                                                # columns equal to iteration number
nmk<-numeric(samp)                              # storage vector also equal to sample size

for( i in 1:100) # bootstrap starts (runs for 10000) iterations
	{ 

	dist1<-rnegbin(samp, dmean, k)      # creates a negbin distribution with your data
	dizzy<-sort(dist1,decreasing = TRUE) #sorts distribution highest parasite count first

		for (j in 1:samp) # a nested loop which counts through the number of hosts

			{

		            addup<-sum(dizzy[1:j])                      # cumulative sum of parasites 
     			    if (sum(dizzy)>0){nmk[j]<-addup/sum(dizzy)} # if there are ANY parasites 
		                                                      # this gives the cumulative   
		                                                      # proportion transmission per 
		                                                      # host
            		    else {nmk[j]<-0 }  # if there are no parasites pop stops 
		                                   # a problem due to division by 0

			}

	mk[,i]<-nmk # puts nmk into matrix mk in column i

	}

fk<-numeric(samp)

for (i in 1:samp)

	{fk[i]<-mean(mk[i,])} # take average of 10000 iterations for each row (each sample)

raw<-data.frame(fk)
props<-c(0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,0.5,
         0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99)
###The proportion of hosts to which we want to fit parasite populations

tvals<-matrix(NA,nrow=dim(raw)[2],ncol=length(props))

for (j in 1:length(props))
	{
	for (i in 1:dim(raw)[2]) 
		{
		vec<-raw[,i]          
		vec<-na.omit(vec)     
		N<-length(vec)        
		Nx<-round(props[j]*N)
		
		if (Nx==0) tvals[i,j]<-0 
		else {tx<-vec[Nx]        
		tvals[i,j]<-tx           
			}
		}
	}

t.vals_Estimated<-data.frame(tvals) 
```

###Data comparison
These distributions can be put together for all the data with complete raw parasite counts. These can then be compared statistically. The data are provided in Appendix S1; Table S2.

```{r}
dat<-read.csv("C:\\Malaria\\SherrardSmithetalGROUNDTRUTH.csv")
stackT20<-c(dat$estimated0.2,dat$real0.2)
stackT40<-c(dat$estimated0.4,dat$real0.4)
stackT60<-c(dat$estimated0.6,dat$real0.6)
stackT80<-c(dat$estimated0.8,dat$real0.8)

##create a data frame with the stacked data
datatype<-c(rep("estimate",27),rep("real",27))
mat<-data.frame(datatype,stackT20,stackT40,
	stackT60,stackT80)

###Compare observed and estimated data
modt20<-glm(log(mat$stackT20)~mat$datatype,family=gaussian) 
summary.lm(modt20)
```

```{r}
modt40<-glm(log(mat$stackT40)~mat$datatype,family=gaussian)
summary.lm(modt40)

modt60<-glm(log(mat$stackT60)~mat$datatype,family=gaussian)
summary.lm(modt60)

modt80<-glm(log(mat$stackT80)~mat$datatype,family=gaussian)
summary.lm(modt80)
par(mfrow=c(2,2))
plot(modt80)
boxplot(log(mat$stackT20)~mat$datatype,ylab="T20")
boxplot(log(mat$stackT40)~mat$datatype,ylab="T40")
boxplot(log(mat$stackT60)~mat$datatype,ylab="T60")
boxplot(log(mat$stackT80)~mat$datatype,ylab="T80")
```


###Step 3
###Calculating Poulin's D and Gini Coefficient
```{r}
##############################################
##
## Poulin's D function (Sherrard-Smith, R version 3.3.3)
## 
##############################################

poulD <- function(dat){
    	N=length(dat)
x=sumx=numeric(N)
  
 	for(j in 1:N){
    		x[j]=mean(dat[1:j])
 		 }
  	for(i in 1:N){
    		sumx[i]=sum(x[1:i])
  		}
PoulinsD = 1 - (2*sum(sumx)/ mean(dat)*N*(N+1))
  	return(list(PoulinsD))
}
##############################################
##
## Gini coefficient function (Sherrard-Smith, R version 3.3.3)
## 
##############################################
Gini <- function(dat){
    	N=length(dat)
	abs_diff = array(dim=c(N,N))
x=sumx=ABS_3=numeric(N) 

	for(j in 1:N){
    		x[j]=mean(dat[1:j])
 		 }
  	for(i in 1:N){
    for(j in 1:N){
      		abs_diff[j,i] = (dat[i] - dat[j]) 
 		 }
         } 
abs_diff2 = ifelse(abs_diff < 0, abs_diff*-1,abs_diff) 
GiniCoef = sum(abs_diff2) / (2 * N * sum(dat))
return(list(GiniCoef))
}
```

###Step 4
###Ecological Analysis
This section provides the script and output of the analyses of the ecological factors potentially important in
determining parasite aggregation in host populations. The data used for this analysis are in Appendix S1; Table S1. This file contains the following columns (for more details please see the
main text):

• **label**: Identifies individual record - these match up with those given in the summary of the references

• **lab.field**: Whether the study was conducted in a lab or field setting

• **parasite.taxa**; parasite.type, parasite.species: Details about the parasites included in the database

• **host.taxa**; host.common.name; host.species: Details about the hosts included in the database

• **transmission**: The parasite transmission mode

• **host.management**: Whether and how the host population is managed

• **habitat**: The host’s habitat

• **eat.meat**: Whether or not the host eats other animals

• **social**: Whether or not the host is social

• **n**: The number of hosts sampled

• **k**: The negative binomial shape parameter

• **mean**: The mean abundance (i.e. total.parasites divided by n)

• **total.parasites**: The total number of parasites recorded in the study

• **gini**: The gini coefficient

• **poulinD**: Poulin’s index of dispersion, D

• **t1-t99**: The proportion of parasites in x proportion of hosts

```{r}
rm(list=ls()) #remove previous objects and packages
dframe1<-read.csv('C://Malaria//SherrardSmithetalDATA.csv')
require(lme4)
require(car)
require(MASS)
require(MuMIn)
require(arm)
require(multcomp)
require(visreg)
require(glmmADMB)
require(ResourceSelection)
```

```{r}
# Remove one study with outlying mean abundance
df2 <- subset(dframe1, label != "227")
```

First, we assessed the correlations between our putative explanatory variables to check their suitability for
inclusion in the model.
```{r}
# This function is from 'Mixed effects models and extensions in
# ecology with R'. (2009).Zuur, AF, Ieno, EN, Walker, N, Saveliev,
# AA, and Smith, GM. Springer.

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y))
txt <- format(c(r, 0.123456789), digits = digits)[1]
txt <- paste(prefix, txt, sep = "")
if (missing(cex.cor))
cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~n + mean + poulinD + gini + t20 + k, data = df2, lower.panel = panel.smooth,
upper.panel = panel.cor, na.action = na.omit)
```

```{r}
mod <- glmmadmb(gini ~ n + log(mean) + transmission + social +
habitat + eat.meat + host.management + parasite.type +
(1 | parasite.taxa/parasite.species) + (1 | host.taxa/host.species),
data = df2, family = "beta")
Anova(mod)
summary(mod)
hist(resid(mod))
```

```{r}
# this function tests for overdispersion. It's from
# http://glmm.wikidot.com/faq
overdisp_fun <- function(model) {
## number of variance parameters in an n-by-n
## variance-covariance matrix
vpars <- function(m) {
nrow(m) * (nrow(m) + 1)/2
}
model.df <- sum(sapply(VarCorr(model), vpars)) + length(fixef(model))
rdf <- nrow(model.frame(model)) - model.df
rp <- residuals(model, type = "pearson")
Pearson.chisq <- sum(rp^2)
prat <- Pearson.chisq/rdf
pval <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
c(chisq = Pearson.chisq, ratio = prat, rdf = rdf, p = pval)
}
overdisp_fun(mod)
```

Hosmer-Lemeshow goodness of fit test with

```{r}
# ResourceSelection package
hoslem.test(df2$gini, y = fitted(mod))
```

###Post-hoc tests of significant effects
####Social structure
Solitary animals have more highly aggregated parasites than social animals.
```{r}
summary(df2$social)
visreg(mod, 'social') # Gives the partial residuals used to plot
# Fig 3.
summary(glht(mod, mcp(social = "Tukey")))
```

####Host habitat
Aquatic parasites are significantly less aggregated than terrestrial ones.
```{r}
summary(df2$habitat)
visreg(mod, 'habitat') #Gives the partial residuals used to plot
# Fig 3.
summary(glht(mod, mcp(habitat = "Tukey")))
```

####Sample size, n
Aggregation increases as you sample more hosts.
```{r}
visreg(mod, "n")
```

