---
title: "Appendix S2: Ecological drivers of parasite aggregation: a meta-analysis of adherence to the 20/80 rule"
author: "E. Sherrard-Smith, J. F. Stephenson, E. Gillingham, R. Brackley, D. Shaw, S. E. Perkins"
date: "26 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This supplement provides the script for the data preparation and calculation of *k*, Gini coefficient, Poulin's D and *t<sub>20*. The script and output of the analyses of the ecological factors potentially important in
determining parasite aggregation in host populations. The data used for this analysis are in the supplementary
file ‘SherrardSmithetalDATA.csv’ (Appendix S1; Table S1). 



## A step by step run through of the data analysis for 20:80 #####

We explore whether parasite populations fit the '20:80' distribution where 80% of the parasites are contained within 20% of the hosts. First, we explore real data on this relationship and then compare these 
distributions to those that are generated using the mean, N of hosts, and a measure of variance.


###Step 1
```{r}
library(MASS)
library(reshape2)
```

###Example Using Real Data; supplied by Darren Shaw September 2013
```{r}
CnodosMarWilliams63 <- rep(seq(0,5,1),c(496,58,18,7,2,4))
```

First data is summarised to give the appropriate information on mean, variance, host and parasite number for the comparison of these real distributions with those that are estimated.

```{r}
data<-sort(CnodosMarWilliams63,decreasing=TRUE)
	data
	sampmean<-mean(data)
	sampmean               ###3.446
		a<-numeric(10000)
		for(i in 1:10000) a[i]<-mean(sample(data,replace=T),na.rm=T)
		quantile(a,c(0.025,0.975))
	sumNpar<-sum(data)
	sumNpar
	N<-length(data)
	N
	var<-var(data)
	var
	k<-(sampmean^2-(var/N))/(var-sampmean)
	k
```


Generate the distribution of parasites in a given proportion of the hosts Using the real data distributions
```{r}
real2<-data.frame(data)

##distribution of parasites in the sample population
real3<-data.frame(sort(real2[,1],decreasing=TRUE))

samp<-length(real3[,1])
nmk<-numeric(samp)
mk<-matrix(data = NA, nrow = samp, ncol = 1) 

for (j in 1:samp) ## a nested loop which counts through the number of hosts
  {
            addup<-sum(real3[1:j,1])             ## cumulative sum of parasites 
     if (sum(real3)>0){nmk[j]<-addup/sum(real3)} ## if there are ANY parasites this gives the cumulative proportion transmission per host
            else {nmk[j]<-0 }                    ## if there are no parasites pop stops a problem due to division by 0
  }

mk[,1]<-nmk                                      ## puts nmk into matrix mk in column i
raw<-data.frame(mk)
props<-c(0.01,0.05,0.1,0.15,0.2,0.25,0.3,0.35,0.4,0.45,0.5,0.55,0.6,0.65,0.7,0.75,0.8,0.85,0.9,0.95,0.99)#the proportions we want
```

```{r}
names<-paste("t",props,sep="")#names for each list of proportions
tvals<-matrix(NA,nrow=dim(raw)[2],ncol=length(props))# a matrix to store the tx values in - rows = studies, cols = tx values


for (j in 1:length(props))# first loop cycles through desired tx values contained in props above
{
	for (i in 1:dim(raw)[2]) #second loop cycles through each column of the data (which is a vector of proportions of parasites in each host from most to least infected)
	{
	vec<-raw[,i]             #make a vector from the column
	vec<-na.omit(vec)        #take out NAs after values reach 1 (it's a ragged array)
	N<-length(vec)           #number of hosts = number of rows
	Nx<-round(props[j]*N)    # gets the row which is x of the way down the column
	if (Nx==0) tvals[i,j]<-0 #if it rounds Nx to 0, there is no such thing as a zeroth row, so make tx 0
	else {tx<-vec[Nx]        # gets the value in that column - the proportion of parasites captured in the x most infected hosts
	tvals[i,j]<-tx           # rowi=study index colj=tx index}

		}
	}

}

#t.vals<-cast(tvals)

##########
#data<-sort(Pkell.place[,1:2],decreasing=TRUE)# copy the data matrix here and change number col to fit data)
#	data

#t.vals[2,]<-t.vals[1,]*sum(data)
#t.vals[2,]

#t.vals<-cast(t.vals)
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
